Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-14 surfaces-write-up/surface-write-up.tex
various edits to surfaces

\ssec{Notation}
\begin{convention}
Write
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$.
\end{convention}

@@

\section{Canonical Rings on Projective Space}
In this section, we bound the degrees of generators and relations for divisors on $\bp^m$, for all $m \geq 1$. The $n = 1$ case was covered in \cite{dorney:canonical}.

@@

For the remainder of this section, we shall fix $m \geq 1$ and choose an isomorphism $\bp^m \cong \proj V$ so that $x_0,\ldots, x_m$ form a basis for $V$. Through the rest of this section, we will think of $x_i$ as a rational section of $H^0(\bp^m, \sco_{\bp^m})$ of degree 1.

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$ and $\deg D_i = \bida_i$. Let $f_i$ be Cartier divisors such that $D_i = V(f_i)$. We shall further assume that among there exist functions $f_0,\ldots, f_{m+1}$ form a basis for degree one rational functions on $\bp^m$.

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ so that 
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot \bida_i \cdot f_i = 0 && \text{ and } &&c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}

\noindent
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=1}^n f_i^{c_i} $ such that
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor && \text{ for } && i > m
\end{align}

>>>>

\ssec{Notation}
\begin{convention}
Write
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$. \todo{Peter: What are the $D_i$'s.}
\end{convention}

@@

\section{Canonical Rings on Projective Space}
In this section, we bound the degrees of generators and relations for divisors on $\bp^m$, for all $m \geq 1$. The $\bp^1$ case was covered in \cite{dorney:canonical}.

@@

For the remainder of this section, we shall fix $m \geq 1$ and choose an isomorphism $\bp^m \cong \proj V$ so that $x_0,\ldots, x_m$ form a basis for $V$.  We will think of $x_i$ as a rational section of $H^0(\bp^m, \sco_{\bp^m})$ of degree 1.

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$ and $\deg D_i = \bida_i$. Let $f_i$ be Cartier divisors such that $D_i = V(f_i)$. We shall further assume that (after reordering) $f_0,\ldots, f_{m}$ form a basis for degree one rational functions on $\bp^m$. \todo{Peter: Make reference to ghost points here?}
\todo{Peter: Also, this is a poorly placed statement of our assumption, because we are not assuming that for the one point/induction case that follows right?}

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot \bida_i \cdot f_i = 0 && \text{ and } &&c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}

\noindent
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor && \text{ for } && i > m
\end{align}
